---
title: "Group 5_Finding Minimum Cost and Maximum Profit Using R"
output: pdf_document
date: "2026-04-27"
---


## PART 1

1. Use the data set assigned to your group  
2. Display the dataset  
3. Check the structure of the data  
```{r}
coffee_dataset_group_5 <- read.csv("coffee_dataset_group_5.csv")
coffee_dataset_group_5
```
```{r}
str(coffee_dataset_group_5)
```
### QUESTION SET 1

How many observations are in the dataset?  
- There are 9 observations.
What variables are included?  
- There are 4 variables in the dataset and all are integers.


## PART 2

1. Create the following new variables:
  - Total Cost
  - Revenue
  - Profit
  - Average Cost
```{r}
coffee_dataset_group_5$Total_Cost <- coffee_dataset_group_5$Fixed_Cost + (coffee_dataset_group_5$Variable_Cost_per_Unit * coffee_dataset_group_5$Quantity)

coffee_dataset_group_5$Revenue <- coffee_dataset_group_5$Quantity * coffee_dataset_group_5$Price

coffee_dataset_group_5$Profit <- coffee_dataset_group_5$Revenue - coffee_dataset_group_5$Total_Cost

coffee_dataset_group_5$Average_Cost <- ifelse(coffee_dataset_group_5$Quantity == 0, NA, coffee_dataset_group_5$Total_Cost / coffee_dataset_group_5$Quantity)
```


## PART 3

Identify:  
  1. The minimum average cost  
  2. The quantity where it occurs
```{r}
min_cost <- min(coffee_dataset_group_5$Average_Cost, na.rm = TRUE)

qty_min_cost <- coffee_dataset_group_5$Quantity[which.min(coffee_dataset_group_5$Average_Cost)]

cat("Minimum average cost is", min_cost, "and it occurs at quantity", qty_min_cost)
```
### QUESTION SET 2

What is the minimum average cost?  
- The minimum average cost is 100.25.

At what quantity does it occur?  
- The minimum average cost occurs in quantity 80.

Why might costs increase after this point?  
- Cost increases after a certain point due to the law of diminishing returns, where adding more inputs results in less efficient production.


## PART 4

Identify:  
  1. Maximum profit  
  2. Quantity where profit is highest
```{r}
max_cost <- max(coffee_dataset_group_5$Profit, na.rm = TRUE)

qty_max_cost <- coffee_dataset_group_5$Quantity[which.max(coffee_dataset_group_5$Profit)]

cat("Maximum Profit is", max_cost, "and it occurs at quantity", qty_max_cost)
```
### QUESTION SET 3

What is the maximum profit?  
- The maximum profit is 1300.  
At what quantity is profit maximized?  
- The maximum profit occurs at quantity 100  
What happens to profit after this point?  
- After quantity 100 profit rapidly declines leading to large losses.


## PART 5

Create two graphs using `ggplot2`.  
Graph 1  
  Plot:  
    1. Quantity (x-axis)  
    2. Total Cost and Revenue (two lines)  
Graph 2  
  Plot:  
    1. Quantity (x-axis)  
    2. Profit (line)  
```{r}
library(ggplot2)

# Graph 1
ggplot(coffee_dataset_group_5, aes(x = Quantity)) +
  geom_line(aes(y = Total_Cost, color = "Total Cost"), linewidth = 1.1) +
  geom_point(aes(y = Total_Cost, color = "Total Cost"), size = 2) +
  
  geom_line(aes(y = Revenue, color = "Revenue"), linewidth = 1.1) +
  geom_point(aes(y = Revenue, color = "Revenue"), size = 2) +
  
  labs(
    title = "Total Cost vs Revenue",
    x = "Quantity",
    y = "Amount",
    color = "Legend"
  ) +
  theme_minimal()
```
```{r}
# Graph 2
ggplot(coffee_dataset_group_5, aes(x = Quantity, y = Profit)) +
  geom_line(linewidth = 1.1, color = "blue") +
  geom_point(size = 2, color = "blue") +
  labs(
    title = "Profit vs Quantity",
    x = "Quantity",
    y = "Profit"
  ) +
  theme_minimal()
```
### QUESTION SET 4

Where do Revenue and Cost intersect?  
- They intersect on two points somewhere around (Total Cost 4940, Revenue 4600) and (Total Cost 13220, Revenue 13800).  
What does that point represent?  
- These points represents the break-even points.



